Prelim Outline

### 2019-06-12 draft.tex
finished preliminary work, 5.5 pages

\documentclass{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{mathtools}

@@

The DMD procedure allows us to develop low-energy effective theories
for physical systems in a systematic manner beginning with a higher energy Hamiltonian H on a Hilbert space $\mathcal{H}$. When trying to develop effective theories for realistic systems H is usually the \textit{ab-initio} Hamiltonian
\begin{equation}
H_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|} 
\label{ab}
\end{equation} 
where the sum $i$ is over electrons and $I$ over nuclei, and the unit for energy is Hartree (Ha). Typically we work under the Born-Oppenheimer approximation and use pseudo-potentials as an efficient replacement for core electrons. This is not always the case, for example when downfolding a second quantized model like shown in equation \ref{hubbard}. The core idea behind the DMD procedure is functional fitting, in particular approximating the total energy functional on a particular subspace of the full Hilbert space. Mathematically:

\begin{equation}
(E[\Psi] \equiv \langle \Psi | H |\Psi \rangle, \mathcal{H}) \rightarrow (E_{eff}[\Psi], \mathcal{LE})\text{ s.t. } E_{eff}[\Psi] = E[\Psi] + \epsilon[\Psi] \ \forall \ |\Psi\rangle \in \mathcal{LE}.
\label{DMD}
\end{equation}
where $\epsilon$ represents an error within our effective theory. The fitting procedure is conducted similarly to fitting any function: 1) Choose a $\mathcal{LE}$ which is spanned by the N lowest energy eigenvectors within $\mathcal{H}$, 2) sample the space to get a sample set $\mathcal{SS} \in \mathcal{LE}$, 3) construct a set of potential models with some variable parameters $\mathcal{MS}$, and 4) regress these models using your sampled data to fit the variable parameters. The density-matrix part of DMD comes from the particular choice of functional parameterization: 
\begin{equation}
E_{eff}[\Psi] = \sum_i c_i d_i[\Psi],\ d_i[\Psi] = \langle \Psi |\hat{d}_i|\Psi \rangle
\end{equation}
where $\hat{d}_i$ is a Hermitian operator in our Hilbert space and many times chosen to be a (sum of) 1- or 2-body density matrices. Common choices of $\hat{d}_i$ are $n_k,\ c_j^\dagger c_k + h.c.,\ n_{k\uparrow}n_{k\downarrow},\ \vec{S}_k \cdot \vec{S}_l,\ $ etc. With this parameterization one can even extract an effective Hamiltonian from the effective energy functional: $H_{eff} = \sum_i c_i \hat{d}_i$ which can be solved and studied. Mapping a downfolding problem onto a regression problem allows us to develop methodologies and systems for building effective theories. In addition we can quantiatively validate our proposed effective theories using tools from data science and statistics. Not only is this method pleasing to statisticians, it is in principle exact. Suppose we want to build a model on $\mathcal{LE}$ composed of the N lowest-energy eigenstates. Let us consider a potential model $H_{eff} = \sum_{i=1}^{ N} c_i |\Psi_i\rangle \langle \Psi_i|$ where the sum is over the N lowest eigenstates with variable coefficients $\{c\}$. If we select \textit{any} linearly independent set of N samples from $\mathcal{LE}$, eigenstates are not required, an ordinary linear regression will lead to the optimal coefficients $c_i = E_i$, the eigenvalues of the eigenvectors $|\Psi_i\rangle$, and our effective model will be exact. In practice we do not know all the eigenstates of H so identifying and sampling $\mathcal{LE}$ is of primary concern, especially for strongly correlated electron systems where the low-energy degrees of freedom may be very complicated collective excitations. I will discuss one approach we had to sampling the low-energy space when talking about our model for the CuO molecule later on. If we do not know the eigenstates we also need to choose descriptors $d_i$ which can compactly represent the variation in properties within $\mathcal{LE}$. Typically one will collect all reasonable $d_i$ like occupation, hopping, exchange coupling, on-site Coulomb repulsion, Hund's coupling, etc ad consider only those $d_i$ which vary within $\mathcal{SS}$. One can then use tools from data science like principal component analysis, step-wise AIC/BIC, regularized regression (e.g. LASSO) for model selection and fitting.

When downfolding \textit{ab-initio} Hamiltonians we will use fixed node diffusion Monte Carlo (FN-DMC) to generate wave functions in our chosen low-energy subspace and to accurately calculate energies and reduced density matrices. Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function. Consider a trial wave function $|\psi\rangle$ and a Hamiltonian $H$ with ground state $|\phi_0\rangle$. We apply the projector $e^{-\tau (H-E_0)}$ as $\tau \rightarrow \infty$ to $|\psi \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau (H-E_0)} |\psi\rangle \propto \langle \phi_0|\psi\rangle |\phi_0\rangle,
\end{equation}
projecting out the ground state $|\phi_0\rangle$ of $H$ as long as the trial wave function we choose is not orthogonal to the ground state. The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(H - E_T)} | R^\prime \rangle$. Since the Hamiltonian has two parts, $H = T + V$ a Trotter expansion is used to approximate the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(H - E_T)} | R^\prime \rangle \sim e^{-\tau(V(R) + V(R^\prime))/2} \langle R| e^{-\tau(T - E_T)}|R^\prime \rangle + O(\tau^2) $ which leads to a time step error in the calculation if the time step between moves $\tau$ is too large. This implementation when used on a Hamiltonian with fermions also leads to a fermion sign problem. We deal with this sign problem through a fixed-node approximation, where the nodal surface of the projected wave function is forced to match that of the initial trial wave function. This approximation makes FN-DMC variational, and will only return the exact ground state of $H$ if the nodal surfaces of $|\psi\rangle$ and $|\phi_0\rangle$ are identical. Therefore the FN-DMC calculation has two major systematic errors that have to be dealt with, the fixed-node error and the time-step error. The time-step error is typically handled through a time-step extrapolation, where multiple FN-DMC calculations are conducted at sequentially smaller $\tau$ and the relevant expectation values are extrapolated to $\tau \rightarrow 0$ \cite{Needs2010}. Fixed-node error can be reduced by using optimized many-body trial wave functions of the (multi-)Slater-Jastrow form seen in \cite{PhysRevLett.98.110201}. We can use the final FN-DMC state to calculate the expectation values of operators like the Hamiltonian or reduced density matrices. For an operator $\hat{O}$ we will use a mixed estimator to calculate the expectation value $\langle \Psi_{DMC} |\hat{O} | \Psi_T \rangle/\langle \Psi_{DMC} | \Psi_T \rangle$ where $\Psi_{DMC}$ is the projected FN-DMC state starting with the trial wave function $\Psi_T$, as this estimator allows us to use the importance sampled distribution $f = \Psi_{DMC}\Psi_T$ for our Monte Carlo estimates. For operators which do not commute with the Hamiltonian H the mixed estimator has an error which can be corrected to first order by subtracting twice the VMC estimate of the expectation value \cite{ceperley_kalos_1979}. Wagner provides the details for calculating reduced density matrix elements in QMC \cite{doi:10.1063/1.4793531}.
\end{document}

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{wrapfig}

@@

The DMD procedure allows us to develop low-energy effective theories for physical systems in a systematic manner beginning with the \textit{ab-initio} Hamiltonian:
\begin{equation}
\hat{H}_\text{ab} = -\frac{1}{2} \sum_{i} \nabla_i^2 - \sum_{i,I}\frac{Z}{|r_i - R_I|} + \sum_{i<j}\frac{1}{|r_i - r_j|}
\label{eq:Hab}
\end{equation}
where $i$ indicates electrons and $I$ nuclei, and the unit of energy is Hartree (Ha).
In this paper we will be working under the Born-Oppenheimer approximation and core electrons will be treated using pseudo-potentials.
The procedure defines a method for downfolding the \textit{ab-initio} energy functional and Hilbert space onto an effective energy functional defined on a low-energy subspace of the Hilbert space 
\begin{equation}
(E_\text{ab}, \mathcal{H}) \xrightarrow{\text{DMD}} (E_\text{eff}, \mathcal{LE} \subset \mathcal{H}), \text{ s.t. }
\forall |\Psi\rangle \in \mathcal{LE}, \ E_\text{eff}[\Psi] - E_\text{ab}[\Psi] = \epsilon[\Psi] + E_0.
\label{eq:DMD}
\end{equation} 
The energy functionals can be written as expectations over Hamiltonian operators $E_\text{ab} = \langle \Psi | \hat{H}_\text{ab} |\Psi \rangle$, $E_\text{eff} = \langle \Psi | \hat{H}_\text{eff} |\Psi \rangle$, $\epsilon$ represents the error in our effective theory, and $E_0$ is a constant energy shift.
Broadly the DMD method consists of four steps which will be discussed in detail later: 
A) Defining the space of low-energy excitations which $E_\text{eff}$ should be able to describe accurately.
B) Sampling, at minimum, a set of states in $\mathcal{H}$ which probe the chosen low-energy excitations in a statistically independent manner. The selected states need not be eigenstates. 
C) Constructing a set of candidate models which are linear in reduced density matrix (RDM) descriptors with variable coefficients $\{c\}$
\begin{equation}
E_\text{eff} = \sum_k c_k \langle \Psi | \hat{d}_k |\Psi \rangle + E_0.
\label{eq:Eeff}
\end{equation}
Here $\hat{d}_k$ are Hermitian operators typically composed of 1- or 2-RDM elements. 
Examples of possible $\hat{d}_k$ are number operators $\hat{n}_i$ and exchange operators $\vec{S}_i \cdot \vec{S}_j$.
Additionally, a single particle basis must be built to express RDM elements on and should be chosen such that variations among states in $\mathcal{LE}$ are mostly described by variations in the projection of these states onto the basis.
D) Using the samples to fit the coefficients $\{c\}$ by linear regression such that $E_\text{eff}$ satisfies the conditions in \eqref{eq:DMD} with the smallest error $\epsilon$ possible. 
The fitting procedure is typically an ordinary linear regression, however, one may use a custom cost function in the regression as well. 
In order to conduct the linear regression for a functional like \eqref{eq:Eeff}, one needs to calculate for each state sampled from $\mathcal{LE}$ the expectation values of $\hat{H}_\text{ab}$ and every descriptor $\{\hat{d}_k\}$ used in the regression.
In principle, if conducted with no approximations, the DMD method is rigorously guaranteed to obtain exact results, namely $\epsilon[\Psi] = 0$.

When downfolding $\hat{H}_\text{ab}$ we will use FN-DMC to access states in $\mathcal{LE}$ and accurately calculate their \textit{ab-initio} energies and RDM elements.
Diffusion Monte Carlo (DMC) is a quantum Monte Carlo method which projects out the ground state of a Hamiltonian given some initial trial wave function.
Consider a trial wave function $|\Psi_T\rangle$ and the Hamiltonian $\hat{H}_\text{ab}$ with ground state $|\Phi_0\rangle$. We apply the projector $e^{-\tau \hat{H}_\text{ab}}$ as $\tau \rightarrow \infty$ to $|\Psi_T \rangle$
\begin{equation}
\lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} |\Psi_T\rangle 
\equiv \lim_{\tau \rightarrow \infty} |\Psi_\text{DMC}(\tau)\rangle \propto \langle \Phi_0|\Psi_T\rangle |\Phi_0\rangle,
\end{equation}
projecting out the ground state as long as the trial wave function we choose is not orthogonal to the ground state. 
The stochastic implementation involves moving samples from the trial function $\Psi_T(R)$ using the Green function $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H}_\text{ab} - E_T)} | R^\prime \rangle$. Since $H_\text{ab} = T + V$, kinetic and potential energy terms, the Green function is approximated by a Trotter expansion $G(R, R^\prime, \tau) = \langle R | e^{-\tau(\hat{H} - E_T)} | R^\prime \rangle \sim \Big[e^{-d\tau(V(R) + V(R^\prime))/2} \langle R| e^{-d\tau(\hat{T} - E_T)}|R^\prime \rangle + O(d\tau^2) \Big]^N $ where $d\tau = \tau/N$.
This expansion can be interpreted as an interative procedure where samples are moved N times with a small timestep $d\tau$ until convergence.
The constant $E_T$ is a trial energy used to control the normalization of $\Psi_\text{DMC}(\tau, R)$ and is updated at each move.
This implementation, however, suffers from a fermion sign problem. 
We deal with this sign problem through a fixed-node approximation, where the nodal surface of $\Psi_\text{DMC}(\tau, R)$ is forced to match that of the initial trial wave function for all $\tau$.
This approximation makes FN-DMC variational, and will only return the exact ground state of $\hat{H}_\text{ab}$ if the nodal surfaces of $|\Psi_T\rangle$ and $|\Phi_0\rangle$ are identical.
We make use of this variational principle to access low-energy states within $\mathcal{H}$ by choosing trial wave functions with nodes that differ from the nodes of $|\Phi_0 \rangle$.
To calculate the expectation value of an operator $\hat{A}$ in FN-DMC we will use a mixed estimator $\langle \Psi_{DMC}(\tau) |\hat{A} | \Psi_T \rangle/\langle \Psi_\text{DMC}(\tau) | \Psi_T \rangle$ which allows us to use the importance sampled distribution $f = \Psi_\text{DMC}(\tau)\Psi_T$ for our Monte Carlo estimates.
Details for calculating reduced density matrix elements in FN-DMC can be found in Wagner.

\pagebreak
\section{Preliminary Work}
\subsection{Non-orthogonal determinants in FN-DMC trial wave functions}
In order to become acquainted with our code QWalk \cite{WAGNER20093390} and QMC algorithms in general I worked on implementing and testing multi-Slater-Jastrow trial functions with optimized non-orthogonal determinants (MSJ+NO) in FN-DMC \cite{Pathak2018}.
We assessed the efficiency and compactness of this new trial function by calculating the ground state energy and single particle densities of a C$_2$ molecule using FN-DMC and comparing to the results when using multi-Slater-Jastrow trial functions with optimized orthogonal determinant trial functions (MSJ+O). 
The workflow involved constructing the un-optimized trial wave functions, optimizing the parameters using an energy optimization method \cite{Toulouse2007}, and finally using the optimized trial functions in an FN-DMC calculation. 
We found that the FN-DMC energy calculated using an MSJ+NO trial function with only 24 determinants was lower than the FN-DMC energy using an MSJ+O trial function with 55 determinants. 
Further, the FN-DMC charge density calculated using MSJ+NO trial functions had stronger bonding character than when using MSJ+O trial functions, a reasonable result as introducing correlations into trial functions allows for electrons to avoid each other while still occupying the same bonding region. 
Our results indicated that using non-orthogonal determinants may lead to more compact multi-Slater-Jastrow trial wave functions for small molecules.

\subsection{Effective theory for CuO molecule using DMD}
As a first step towards developing accurate models for extended systems like solids using DMD, we constructed a many-body effective model for the CuO molecule which accurately describes the eigenstates and spectra seen in experiment up to 2eV above the ground state.
The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
While the low-energy excitations in the doublet sector of the CuO molecule are well understood, singlet selection rules for spectroscopic measurements and the doublet ground state leave the quartet sector of excitations unexplored.
We construct our $\mathcal{LE}$ based on recent anion photoelectron spectroscopy (APES) measurements of the low-lying excited states of the nuetral CuO molecule. 
The lowest energy excitations measured, which lie below 2eV, involve primarily the O 2p$_z$, 2p$_\pi$ and Cu 4s orbitals, which are fully-filled, half-filled and empty in the ground state respectively.
Beginning at 2.2 eV are excitations which remove a single electron from the fully filled Cu 3d shell and add them to either the O 2p$_\pi$ or Cu 4s orbitals.
The neutral copper atom also carries a Hund's coupling between the d and s orbitals of about $\sim$ 0.5 eV, the effects of which would be measurable among the states in $\mathcal{LE}$ as long as states of both doublet and quartet states are included.
In order to probe all of these low-energy excitations we select a low-energy space
\begin{equation}
\mathcal{LE} = \text{Span(}\{ |\Psi \rangle | \langle \Psi | \hat{n}_{3d} | \Psi \rangle \ge 9,\ \hat{H}_\text{ab}|\Psi\rangle = E |\Psi\rangle \}\text{)}.
\label{eq:LE}
\end{equation}

Our sample space was generated by using multi-Slater-Jastrow trial wavefunctions in FN-DMC. 
In principle we could probe every low-energy excitation in $\mathcal{LE}$ in a statistically independent manner by just sampling the eigenstates within $\mathcal{LE}$ and including various non-eigenstates within $\mathcal{LE}$ as well to ensure robust statistics.
In practice we do not have access to these eigenstates and approximate methods are required to sample the low-energy space. 
Consider a trial wavefunction of the form $e^{J_i} |\text{D}_i\rangle$, where $e^{J_i}$ is a three-body Jastrow factor and $|\text{D}_i\rangle$ is a determinant of some single particle orbitals.
If the nodes of $|\text{D}_i\rangle$ are similar to the nodes of an eigenstate $|\Phi_i\rangle \in \mathcal{LE}$, then the properties and energy calculated from FN-DMC using this trial wavefunction will approximate those of the eigenstate, with the degree of approximation depending on the quality of the trial function nodal structure.
If we were able to find for each $|\Phi_i \rangle \in \mathcal{LE}$ such a $|\text{D}_i\rangle$ then we can approximate any state in $\mathcal{LE}$ as
\begin{equation}
|\Psi\rangle = \sum_i c_i |\Phi_i\rangle \sim \sum_i c_i \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J_i} |D_i\rangle) \xrightarrow{J_i = J} \lim_{\tau \rightarrow \infty} e^{-\tau \hat{H}_\text{ab}} (e^{J}\sum_{i} c_i|\text{D}_i\rangle).
\label{eq:sampling}
\end{equation}
The last equality is valid if $\forall i\  J_i = J$, allowing us to map $|\Psi \rangle \in \mathcal{LE}$ to the end result of an FN-DMC calculation using a multi-Slater-Jastrow trial wavefunction. 
We use symmetry-targeted unrestricted Kohn-Sham (UKS) to generate our determinants $|\text{D}_i\rangle$ using a B3LYP functional, which has been shown to accurately reproduce ground state nodal properties of transition-metal oxide molecules, a Trail-Needs pseudopotential, and the VTZ Trail-Needs basis using the package PySCF.
This method allows us to access almost every excitation in $\mathcal{LE}$, however, there are some cases of two or more low-energy states within $\mathcal{LE}$ which have identical S$_z$ and number of electrons per irrep.
An example of this dilemma would be the ground state and the state with an excitation from Cu 3d$_{xz} \rightarrow $  O 2p$_x$.
In these few cases the higher energy excited $|\text{D}_i\rangle$ is inaccessible and the corresponding $|\Phi_i\rangle$ is excluded from the sum in \eqref{eq:sampling}, restricting our sample space to a subspace of $\mathcal{LE}$.
We will discuss the consequences of this restriction later on.
A single three-body Jastrow factor $J$ was used for every $|\text{D}_i\rangle$ and was optimized on the lowest energy UKS state using a linear energy optimization method.
The FN-DMC calculations were conducted with T-moves, to make the calculation variational while using pseudopotentials, with a timestep of $\tau = 0.01$.
The coefficients $\{c_j\}$ for our sample states were chosen via a shell-sampling method. 
We begin by fixing the coefficient $c_0 = \sqrt{w}$ where $w \in \{1.0, 0.8, ..., 0.2\}$. 
For each choice of $w$ we sample n = 5 states by randomly selecting the unassigned coefficients from a uniform distribution such that $\sum_i c_i^2 = 1$. 
This procedure generates shells of samples at decreasing distance from the eigenstate $|\Phi_0\rangle$. 
We can then loop over all $i$ to generate shells near each eigenstate.

Parameterizing our effective model requires constructing a basis on which to calculate our \textit{ab-initio} 1-/2-RDM elements and selecting a set of descriptors from which we would like to build our model.
We constructed two such bases: a localized intrinsic atomic orbital (IAO) basis and a molecular orbital (MO) basis.
The IAO basis was built on all Cu 3d, 4s and O 2p molecular orbitals from the $|\text{D}_i\rangle$ used to construct our sample states.
The MO basis consists of the Cu 3d, 4s and O 2p molecular orbitals from the lowest energy restricted open-shell Kohn-Sham (ROKS) solution with $x \leftrightarrow y$ spatial symmetry, a O 2p$_z \rightarrow$ O 2p$_\pi$ excitation.
As these two bases are used to describe the same space, they are related by an orthogonal transformation which will allow us to map between them if necessary.
In order to describe the low-energy degrees of freedom in the CuO molecule our model will require occupation energies and hybridizations between the Cu 3d, 4s and O 2p atomic orbitals which are most compactly represented in an MO basis.
The lack of doubly occupied Cu 4s states in the low-energy spectrum of CuO indicate the necessity of a Hubbard U term on the Cu 4s.
This is corroborated by the fact that the FN-DMC energy of the lowest-lying sampled state with a doubly occupied Cu 4s orbital is 7 eV above the lowest energy sampled state.
Further, we include a Hund's coupling between the Cu 4s and Cu 3d orbitals in our model based on the existence of a similar Hund's coupling 
for the copper atom.
These two interaction terms are best represented in the IAO basis as they are intuitively interactions between localized orbitals.
In addition to these important parameters, we will consider models which contain any number of hybridizations between the MOs to account for orbital relaxation between excited states in the CuO molecule.
In total our space of canditate models is then 
\begin{equation}
\begin{split}
\{\bar{n}_{4s}, \bar{n}_{p_\pi}, \bar{n}_{p_z}, \hat{n}_{4s\uparrow} \hat{n}_{4s\downarrow},\sum_{i \in \{xy, xz, ...\}}\vec{S}_{4s}\cdot \vec{S}_{d_i}\} \ + \\
\text{P}(\bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{4s}^\dagger \bar{c}_{p_z} + h.c.,\ \bar{c}_{d_z^2}^\dagger \bar{c}_{4s} + h.c.)
\end{split}
\label{eq:models}
\end{equation}
where P denotes the power set and operators are defined on the IAO basis unless denoted by a bar, in which case the operator is built on the MO basis. The coefficients for each term above will be denoted $\bar{\epsilon}_{4s},\ \bar{\epsilon}_{p_\pi},\ \bar{\epsilon}_{p_z},\ U_s,\ J_{sd},\ \bar{t}_\pi,\ \bar{t}_{dz},\ \bar{t}_{sz},\ \bar{t}_{ds}$. The operator $\bar{n}_{3d}$ is not included in any models as it is linearly dependent on all other occupation energies through the relationship $\bar{n}_{3d} + \bar{n}_{p_z} + \bar{n}_{p_\pi} + \bar{n}_{4s} = N_\text{elec} = \text{15}$. The symbol $\pi$ denotes a contraction over $x, y$ for p orbitals and $xz,\ yz$ for d orbitals. The symbol $\delta$, introduced later, denotes a contraction over $xy,\ x^2-y^2$ for d orbitals.

After fitting each of the potential models in \eqref{eq:models} using ordinary linear regression (OLS) and solving the resultant models using exact diagonalization, we find a set of models which describe the energy functional on our sample set accurately but whose eigenstates and spectra differ as a consequence of intruder states below an energy of 2eV. 
The sixteen different models all regress with R$^2 >$ 0.95 on our sample data, however some models are more robust than others. 
Models which have coefficients that are zero within 95\% confidence intervals, computed through bootstrap estimates, are removed from consideration.
Any models which have an average errorbar in the first 20 exactly diagonalized eigenvalues, also estimated by bootstrapping, greater than 0.20 eV are ignored.
Each of the remaining three models describe our sample set accurately, have robust non-zero coefficients, and have small error bars in eigenvalues of states which we care about the most.
However, the eigenstates below 2eV differ greatly between these models as illustrated in Figure \ref{fig:InitED}.

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/init_ed.pdf}
\end{center}
\caption{Results of exact diagonalization showing the energies and properties of the first twenty eigenstates for three candidate models fit using ordinary linear regression. The ground state and first excited state are colored and have enlarged markers for clarity. Error bars are 95\% CI calculated by a bootstrap estimate.}
\label{fig:InitED}
\end{figure}

All three models have a first excited state around 1eV, but only in the minimal model does this state fall near the sample set used for the fitting, and is the only model in which this state corresponds to the correct excitation O 2$p_z \rightarrow $ O 2p$_\pi$ as seen in experiment.
In contrast, the first excited state for the other two models are visibly distant from our sample set and do not correspond to the experimental first excited state.
The large distance of these eigenstates from our sample set indicates that their predicted eigenvalues may suffer from large extrapolation errors, and as such we will call them \textit{intruder} states.
In Figure \ref{fig:Intruder} we present all eigenstates among the three potential models which we classify as intruders using a k-nearest neighbor approach with k = 5.

Given that our intruder states correspond to states within $\mathcal{LE}$ but outside our sample set, they must live within the span of the eigenstates which were excluded in our sampling scheme as described before.
This is evidenced by the fact that two of the intruder states correspond explicitly to excitations we could not access: Cu 3d$_{z^2} \rightarrow $ O 2p$_\pi$ and  Cu 3d$_\pi \rightarrow $ O 2p$_\pi$.
In fact, these are the only two singly excited states which we could not access with our sampling scheme.
While inaccessible, we can approximate these states by single rigid MO excitations above the UKS ground state, which results in states with a FN-DMC energy above 4eV.
Including a generous 2eV reduction in FN-DMC energy due to orbital relaxation, we estimate that these two states should lie at least 2eV in energy. 
Since all the other inaccessible eigenstates are double or higher order excitations, our very approximate exploration of the inaccessible parts of $\mathcal{LE}$ leaves us with a reasonable prior: intruder states should lie above 2eV in energy in our effective theories.
We can enforce this prior by adding a term to our ordinary linear regression cost function
\begin{equation}
\text{Cost} = \sum_{i} (E_\text{eff}[\Psi_i] - E_\text{ab}[\Psi_i])^2 + \lambda \sum_{p}\text{QHL}(2 - E_\text{eff}[\Psi_p]),\ \text{QHL}(x) = \Theta(x)x^2
\label{eq:cost}
\end{equation}
where $\lambda>0$ is a parameter which can be varied, QHL is a quadratic hinge loss, $\Theta$ is the Heaviside step function, the index $i$ is over our sampled states and $p$ over the selected intruder states.
Increasing the value of $\lambda$ leads to a stronger penalty for models which do not satisfy our prior.

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/intruder.pdf}
\end{center}
\caption{Collection of unique intruder states from our three potential models selected using a k-nearest neighbor approach with k = 5.}
\label{fig:Intruder}
\end{figure}

Using our new cost function we find a set of models which accurately describe our sample data, have very similar spectra and eigenvectors, and do not contain intruder states below the 2eV prior. 
Shown in Figure \ref{fig:Prior} are the R$^2$ scores and QHL for the three candidate models fit at $\lambda \in [0,20]$. 
The R$^2$ score stays above 0.95 for all models and $\lambda$, and the QHL converges quickly to nearly zero at $\lambda = 20$ for all three models. 
We conclude that all three models at $\lambda = 20$ should be considered good models as they satisfy our prior and also explain our sample set accurately. 
Also shown are $E_\text{eff}, E_\text{ab}$ for our samples using the minimal model at $\lambda = 20$ to illustrate the high fit quality even with a large $\lambda$.
Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for the three fit models at $\lambda = 20$. 
In comparison one can view Figure \ref{fig:InitED} as the same figure for fits at $\lambda = 0$. 
At $\lambda = 20$ there are no outliers below 2eV in any of the three fit models. 
The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
This conclusion is verified by looking at the values of the fit parameters for our three models, presented in Table 1.
The parameter values for the three different models fit at $\lambda = 20$ are nearly identical within error bars and indicate that asserting our prior ensures that all candidate models regress to a single low-energy effective theory.
All occupation energies are relative to $\epsilon_{d_\delta}$.

Our final regressed model accurately reproduces properties and energies of the lowest lying eigenstates as seen in modern APES experiments on the CuO molecule.
From Figure \ref{fig:FinalED} we see that the ground state of our effective theory has fully filled Cu 3d shell, nearly two electrons in O 2p$_z$ orbital, three in the O 2p$_\pi$ and around a quarter of an electron in the Cu 4s orbital.
An approximate electronic configuration for the ground state is therefore
be assigned 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^3$3$\sigma^0$ where the orbitals associated with the assignment are Cu 3d$_{z^2}$, Cu 3d$_\pi$, Cu 3d$_\delta$, O 2p$_z$, O 2p$_\pi$, and Cu 4s.
This state also has a total spin projection S$_z = \frac{1}{2}$.
This matches the well studied ground state electron configuration for the molecule $^2\Pi$.
There are two excited states around 1eV in energy, the first of which is a state with the following single excitation O 2p$_z \rightarrow$ O 2p$_\pi$, has the configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^4$4$\sigma^0$, and S$_z = \frac{1}{2}$.
According to experiment there is also such a state at 1eV and is usually denoted as $^2X$.
A second excited state at 1.2 eV exists which corresponds to an excitation 
O 2p$_\pi \rightarrow$ Cu 4s, has configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^2$2$\pi^2$4$\sigma^1$, and can either be S$_z = \frac{1}{2}, \frac{3}{2}$. 
This is also the lowest energy eigenstate which can be a spin quartet.
\textbf{ref} agree that this excitation should be the lowest spin quartet excitation, but argue that this state should be at an energy of 1.9 eV. 
Our model has a different spin quartet state with a full Cu d shell at 1.9 eV, corresponding to an excitation O 2p$_z \rightarrow$ Cu 4s, with configuration 1$\sigma^2$1$\pi^4$1$\delta^4$2$\sigma^1$2$\pi^4$4$\sigma^1$. 
There are another host of eigenstates for our models at 2 eV corresponding to singles excitations of the form Cu 3d $\rightarrow$ O 2p$_\pi$. 
This matches measurements which indicate that the lowest energy eigenstates with a single hole in the Cu 3d shell lie at 2.2 eV above the ground state.

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/prior_and_regr.pdf}
\end{center}
\caption{On the left, scores of the three candidate models at various $\lambda$ when fitting our effective theories using the cost function \eqref{eq:cost}. On the right we show predicted versus calculated energies for our low-energy sampled states for the minimal model fit at $\lambda = 20$.}
\label{fig:Prior}
\end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{l|llllll}
&Min & Min & Min \\
& & $ + \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}$ & $ + \bar{c}_{d_{z^2}}^\dagger \bar{c}_{p_z}$ \\ $\ $ \\ \hline 
$\epsilon_{d_{z^2}}$& 0.32(2)& 0.32(1)& 0.29(2)\\
$\epsilon_{d_\pi}$& 0.19(1)& 0.10(2)& 0.19(1)\\
$\epsilon_{d_\delta}$& 0.0 & 0.0 & 0.0\\
$\epsilon_{4s}$& 2.2(4)& 2.23(3)& 2.19(5)\\
$\epsilon_{p_\pi}$& 1.7(3)& 1.8(2)& 1.7(3)\\
$\epsilon_{p_z}$& 0.96(4)& 0.97(3)& 0.99(5)\\
$t_\pi$& -0.57(1)& -0.45(2)& -0.57(1)\\
$t_{dz}$& 0.55(3)& 0.56(2)& 0.54(2)\\
$t_{sz}$& 0.87(2)& 0.89(2)& 0.85(2)\\
$t_{ds}$& 0.44(1)& 0.45(1)& 0.45(2)\\
$J_{sd}$& -0.6(1)& -0.8(1)& -0.6(1)\\
$U_s$& 3.8(2)& 3.7(2)& 3.9(2))\\
\end{tabular}
\\ $\ $
\\
\end{center}
Table 1: Parameters in eV for our three potential models when fit using \eqref{eq:cost} at $\lambda = 20$. Error bars are 95\% CI calculated using a bootstrap estimate.
\end{table}

\begin{figure}[H]
\centering
\includegraphics[width=0.7\linewidth]{../cuo/qwalk/old/ub3lyp_s1_/analysis/figs/final_ed.pdf}
\caption{Results of exact diagonalization showing the energies and properties of the first thirty eigenstates for three candidate models using linear regression with cost function \ref{eq:cost} and $\lambda = 20$. The ground state and first excited state are colored and have enlarged markers for clarity. Error bars are 95\% CI calculated by a bootstrap estimate.}
\label{fig:FinalED}
\end{figure}

\end{document}
